The impact of cosmic variance on simulating weak lensing surveys

### 2014-10-07 Writeup/paper_v3.tex
detailed editing of section 2

 Realistic simulations may use samples based on images from the Hubble
 Space Telescope ({\em HST}).
Software packages like
GalSim\footnote{\url{https://github.com/GalSim-developers/GalSim}}
\citep{2014arXiv1407.7676R} can generate images of galaxies from the
{\em HST} as they would appear with an additional lensing shear and
viewed by some lower resolution telescope.  Examples of training samples from the {\em HST} include 
the COSMOS survey \citep[used by the GREAT3 challenge,][]{great3} or the
 Ultra Deep Field \citep[UDF, used by][]{2013ApJ...765...74J}.  These
 two examples serve as the extremes in the {\em HST} samples used as
 the basis for image simulation, with COSMOS being shallower but
representing the current widest contiguous area surveyed by the {\em HST}, and the
 latter being extremely deep but narrow.

@@

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
The paper is structured as follows: in Sec.~\ref{S:data}, we describe
the data that we use for this study.  In Sec.~\ref{S:methods}, we
describe our methods for deriving the relevant galaxy properties
like environment, morphology, and shape.  
Using these ingredients, we present our results in Sec.~\ref{S:results}
and discuss their implications in Sec.~\ref{S:summary}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
COSMOS~\citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST).

We apply the following set of initial cuts to the data, all of which is defined in~\cite{COSMOS_Alexie}:
\begin{enumerate}
 \item \texttt{MU\_CLASS=1}: Unlike SExtractor's stellar index \texttt{CLASS\_STAR} which is continuous, the \texttt{MU\_MAX} method that compares the peak surface brightness to the background level,
 provides a clear differentiation of galaxies (\texttt{MU\_CLASS=1}) from stars (\texttt{MU\_CLASS=2}) and other spurious objects like cosmic rays (\texttt{MU\_CLASS=3})

 \item \texttt{CLEAN=1}: Objects near bright stars or those containing saturated pixels were removed and the rest classify as \emph{clean}.

% \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
%                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
%                                       \arun{copied from Mandelbaum
%                                         2011.}
 \item \texttt{GOOD\_ZPHOT\_SOURCE =1}: This cut requires that the photometric redshifts be reliable and good enough to draw conclusions about the population. More details on what `good' refers to can be found in \cite{SHERA}
\end{enumerate}
%\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{\url{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/}} for the GREAT3 challenge~\citep{great3}, using the procedure described in \cite{SHERA}.
Parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc were done for GREAT3 challenge~\citep{great3} using the method described in \cite{Claire_Fits}. This is described briefly in Sec.~\ref{sub:axisratio}.In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .

Photometric redshifts were determined by~\cite{COSMOS_Photoz_30band}.The accuracy of photometric redshifts for $m_\text{F814W}\le$ 22.5 is $\sigma_{\Delta z} = 0.007(1+z_s)$ and at fainter magnitudes, i.e. $m_\text{F814W}\le$ 24, $\sigma_{\Delta z} = 0.012(1+z_s)$.
The photometric redshift values become noisy beyond $z$ of $1$ for our purposes and the various fits to the galaxies are also not very reliable beyond the apparent magnitude $(m)$ value of 23.5. 
Thus, we will exclude all galaxies that have F814W magnitude greater than 23.5. However, we will use $m_\text{F814W}\le 25.2$ sample in volume-limiting process (cf.~\S\ref{sub:volumelimiting}) to calculate completeness.  We take the $z<1.0$ flux-limited sample to fit redshift distribution models (cf.~\S\ref{sub:overdensities}) and then restrict ourselves to $z\le1$ sample for all further analysis.
% ($S/N \sim 50$ {\bf ?} )
% \rachel{Our catalog has the more up-to-date Ilbert et al. photo-$z$,
%   not the older Mobasher ones.  Need to fix citations. Also, we care
%   about the photo-$z$ accuracy beyond 22.5, so it's important to give
%   some information about it to the depth that we actually use.  Finally, I
%   recommend against mentioning the $S/N$ because the calculation is
%   complicated by the correlated noise, so most people use use
%   magnitude instead of $S/N$.}

Stellar mass estimates were obtained~\citep{COSMOS_XRAY} using the Bayesian code described in~\cite{KEVIN_MSTAR}. Basically, a grid of models that vary in age, star formation history, dust content and metallicity
are constructed to which the observed galaxy's spectral energy distribution (SED) and the photometric redshift are referenced to. At each grid point, the probability that the 
SED fits the model are calculated and by marginalizing over all the parameters in the grid, the stellar mass probability distribution is obtained. The median of this distribution
is taken as the stellar mass estimate. A Chabrier IMF~\citep{ChabrierIMF} is assumed.

We will analyze how the galaxy morphology and intrinsic shape depends on
the environment in which they reside.

>>>>

 Realistic simulations may use samples based on images from the Hubble
 Space Telescope ({\em HST}).
Software packages like
{\sc GalSim}\footnote{\url{https://github.com/GalSim-developers/GalSim}}
\citep{2014arXiv1407.7676R} can generate images of galaxies from the
{\em HST} as they would appear with an additional lensing shear and
viewed by some lower resolution telescope.  Examples of training samples from the {\em HST} include 
the COSMOS survey \citep[used by the GREAT3 challenge,][]{great3} or the
 Ultra Deep Field \citep[UDF, used by][]{2013ApJ...765...74J}.  These
 two examples serve as the extremes in the {\em HST} samples used as
 the basis for image simulation, with COSMOS being shallower but
representing the current widest contiguous area surveyed by the {\em HST}, and the
 latter being extremely deep but narrow.

@@

%The structure of the paper is as follows. In \S\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
The paper is structured as follows: in Sec.~\ref{S:data}, we describe
the data that we use for this study.  In Sec.~\ref{S:methods}, we
describe our methods for deriving the relevant galaxy properties
like environment, morphology, and shape.  
Using these ingredients, we present our results in Sec.~\ref{S:results}
and discuss their implications in Sec.~\ref{S:summary}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
The COSMOS survey \citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST).  We use the COSMOS survey to
define a parent sample of galaxy images to be used for making image
simulations, following the approach taken to this problem in
\cite{2012MNRAS.420.1518M,great3}.

We apply the following set of initial cuts to the COSMOS data, the
first two of which
are motivated and explained in more detail by \cite{COSMOS_Alexie}:
\begin{enumerate}
 \item \texttt{MU\_CLASS=1}: This criterion uses a comparison between
   the peak surface brightness and the background level to achieve a
   robust star/galaxy separation, with galaxies having \texttt{MU\_CLASS=1}. 
 \item \texttt{CLEAN=1}: Objects near bright stars or those containing
   saturated pixels were removed; the rest pass this cut on \texttt{CLEAN}. 
% \item \texttt{GOOD\_ZPHOT\_SOURCE=1}: This cut requires that there be a good photometric redshift, which typically is equivalent to requiring that the galaxy not be located within
%                                       the masked regions of the Subaru $BVIz$ imaging used for photometric redshifts, and that it have a successful match against an object in the Subaru imaging
%                                       \arun{copied from Mandelbaum
%                                         2011.}
 \item \texttt{GOOD\_ZPHOT\_SOURCE =1}: This cut requires that
   photometric redshifts be reliable and good enough to draw
   conclusions about the population (see \citealt{2012MNRAS.420.1518M}
   for details).
\end{enumerate}
%\rachel{The cuts you just gave are really basic or fundamental ones.  Those should go before any discussion of more complicated derived quantities like redshifts, stellar mass, Sersic fits, etc.}

%Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
%is $0.12''$. 
High resolution images taken through the wide F814W filter (broad
\emph{I}) for all galaxies passing the above cuts were used to create
a collection of postage stamp images for the GREAT3 challenge
\citep{great3}, using the procedure described in
\cite{2012MNRAS.420.1518M}.  Each galaxy postage stamp image has a
corresponding PSF image that can be used by {\sc GalSim} or other
software to remove the effects of the {\em HST} PSF before simulating
the galaxy image as it would appear at lower resolution.

To better characterize the galaxy population, parametric models were
fit to the light profiles of these galaxies.  These were carried out
using the method described in \cite{Claire_Fits}, and include
\sersic{} profile fits and 
2 component bulge $+$ disk fits described in detail in \cite{great3}
and briefly in Sec.~\ref{sub:axisratio} of this work.

In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam, the
Canada-French Hawaii Telescope (CFHT) and KPNO/CTIO, yielding many bands
of imaging data from which to determine high-fidelity photometric
redshifts. 
Photometric redshifts were determined by
\cite{COSMOS_Photoz_30band}. The accuracy of photometric redshifts for
$m_\text{F814W}\le$ 22.5 is $\sigma_{\Delta z} = 0.007(1+z)$; for $m_\text{F814W}\le$ 24, $\sigma_{\Delta z} = 0.012(1+z)$.
The photometric redshift values become noisier beyond $z\sim 1.2$, and
the  fits to the galaxy light profiles are also somewhat noisy once we
go beyond $m_\text{F814W}\sim 23.5$.   
For this reason, we will exclude all galaxies that have F814W
magnitude fainter than 23.5. However, we will use the
$m_\text{F814W}\le 25.2$ sample that was generated for the GREAT3
challenge to estimate the completeness, which is useful when
generating a volume-limited sample (Sec.~\ref{sub:volumelimiting}).
We first use the $z<1$ flux-limited sample to fit parametric redshift
distribution models (Sec.~\ref{sub:overdensities}), and then
restrict ourselves to $z\le1$ sample for all further analysis. 
% ($S/N \sim 50$ {\bf ?} )
% \rachel{Our catalog has the more up-to-date Ilbert et al. photo-$z$,
%   not the older Mobasher ones.  Need to fix citations. Also, we care
%   about the photo-$z$ accuracy beyond 22.5, so it's important to give
%   some information about it to the depth that we actually use.  Finally, I
%   recommend against mentioning the $S/N$ because the calculation is
%   complicated by the correlated noise, so most people use use
%   magnitude instead of $S/N$.}

Stellar mass estimates were obtained~\citep{COSMOS_XRAY} using the
Bayesian code described in~\cite{KEVIN_MSTAR}.  This process involves
constructing a grid of models that vary in age, star formation
history, dust content and metallicity (always assuming a Chabrier IMF;
\citealt{ChabrierIMF}), 
 to which the observed galaxy spectral energy distributions (SEDs) and photometric redshift are compared. At each grid point, the probability that the 
SED fits the model is calculated, and by marginalizing over the nuisance parameters in the grid, the stellar mass probability distribution is obtained. The median of this distribution
is taken as the stellar mass estimate.

%We will analyze how the galaxy morphology and intrinsic shape depends on
%the environment in which they reside.
